First steps towards a formalization of Forcing

### 2018-06-22 LSFA/intro.tex
forcing

The concept of relativization and absoluteness are central to task of
transferring truth of axioms in $M$ to  $L^M$. At this point we 
point out that although $L^M$ is a subset of $M$. 
\bigskip

\hrule
\bigskip

In general, one can relativized

It can be shown elementarily that if $M$ and $N$ are transitive,
formula $\phi$ holds for $x,y,z\in N$

It is noteworthy that

 One of the main
technical tools designed by him

\begin{enumerate}
\item $\CH$ is (Platonistically) undecidable.
  \begin{enumerate}
  \item By G\"odel's Second Incompleteness Theorems, we can't show
    that a model of $\ZFC$ exists.
  \item G\"odel showed that $\CH$ is relatively consistent with
    $\ZFC$. ``Easy.''
  \item Cohen showed that  $\neg\CH$ is relatively consistent with
    $\ZFC$. Not ``easy.'' Formalization Open.
  \end{enumerate}
\item Using $\ZF$, G\"odel defined the constructible universe,
  $L$. G\"odel showed that $\lb L,\in\rb\models \ZFC+\CH$. Paulson formalized $L$
  for the consistency of $\AC$.
\item The main concept for G\"odel's proofs is
  \emph{relativization}. This is the core of Paulson's development.
  \begin{enumerate}
  \item \dots
  \end{enumerate}
  %
\item 
\item Cohen invented Forcing.
  \begin{enumerate}
  \item Countable transitive models $M$. These are enough for consistency
    proofs by the L\"owenheim-Skolem Theorem, the Reflection
    Principle, and the Mostowksi Collapse.
    %
  \item 
    %
  \item Forcing is a technique to extend countable transitive models of
    $\ZFC$. This process is guaranteed to preserve the $\ZFC$
    axioms but allows to fine-tune what other first order properties the
    extension will have. 
    %
  \item Given a preorder with top $\lb\PP,\leq,\1\rb$ in $M$, define
    \emph{filter} $G$, \emph{dense}. In general, there is no  $G\in M$
    that intersects every dense
    subset of $\PP$ that lies in $M$ (``generic'').
  \item Rasiowa-Sikorski states that there is such $G$ for every
    family of countable dense subsets. Thus, there are generic filters
    for countable transitive models.
  \end{enumerate}

\end{enumerate}

>>>>

The concepts of relativization and absoluteness are central to task of
transferring truth of axioms in $M$ to  $L^M$.

\subsection{Forcing}
Forcing is a technique to extend countable transitive models of
$\ZFC$. This process is guaranteed to preserve the $\ZFC$
axioms but allows to fine-tune what other first order properties the
extension will have. Given a ctm $M$ of $\ZFC$, a new set $G$ is
adjoined to $M$ and under some hypothesis ($G$ being ``generic''),
there exists a smallest model $M[G]$ of $\ZFC$ that includes $M$ and
contains $G$.

\fbox{The next should be spread accross the other sections}

The easiest way to define genericity is by using preorders.
Given a preorder with top $\lb\PP,\leq,\1\rb$ in $M$, define
\emph{filter} $G$ ... , \emph{dense} ...  $G$ is an $M$-\emph{generic}
if it intersects every dense subset of $\PP$ that lies in $M$.

But since $M$ is countable,

The  Rasiowa-Sikorski theorem  states that that for every countable 
family of dense subsets there is a filter that intersects them
all. Thus, there are generic filters $G$ for countable transitive models.  In
general, no such  $G$ belongs to $M$ and therefore the extension
$M[G]$ is proper.
